---
title: 'Host-pathogen dynamics in fungal diseases: Comparing SI and multi-infection
  models'
author:
- affiliation:
  - UCD
  email:
  - nmross\@ucdavis.edu
  footnote:
  - Corresponding author
  name:
  - Noam Ross
date: "December 18, 2014"
output:
  html_document:
    dev:
    - svg
    - png
    fig_caption:
    - yes
    fig_height:
    - 6
    fig_width:
    - 12
    keep_md:
    - yes
  pdf_document:
    dev:
    - cairo_pdf
    - cairo_ps
    fig_caption:
    - yes
    fig_height:
    - 6
    fig_width:
    - 12
    keep_tex:
    - yes
    template:
    - elsarticle.latex
csl: ecology-letters.csl
address:
- address:
  - |
    Department of Environmental Science and Policy
    University of California-Davis
    1 Shields Avenue Davis, CA 95616 USA
  code:
  - UCD
documentclass: elsarticle
layout: review,12pt
linenumbers: yes
bibliography: references.bib
abstract: |
  Emerging fungal disease pose major threats to plants and wildlife. As effects of
  many fungal diseases are load-dependent, they may be better represented by
  models with variable loads (multi-infection), rather than susceptible-infected
  (SI) models. I compare equilibrium and transient behavior between these model
  types, including the dynamics of age structure. To compare models with different
  structure, I parameterized multi-infection models to replicate SI behavior at
  either initial or equilibrium phases of an epidemic, comparing the resultant
  behavior at other phases. Multi-infection diseases have either slower initial
  growth rates of disease than SI models, or have greater population-suppressing
  effects. Multi-infection models also produce greater relative mortality in older
  host stages, and greater overall changes in population age structure, than SI
  models. These results indicate that using SI models to predict epidemic behavior
  when diseases are load-dependent may underestimate long-term effects on population
  size.
---

```{r get_chunks, echo=FALSE, results='hide', message=FALSE, warning=FALSE, cache=FALSE}
library(knitr)
read_chunk('base_parms.R')
read_chunk('comparative_simulations.R')
opts_knit$set(eval.after = "fig.cap")
opts_chunk$set(echo=FALSE, cache=TRUE, message=FALSE, warning=FALSE)
```

```{r setup}
<<setparms>>
<<load_pkgs_code>>
```

Introduction
------------

Emerging fungal infections pose major threats to plant and animal wildlife
populations as well as livestock and crops. Fungal life-history traits, including high
virulence, long-lived environmental reservoirs, and host generalism are thought
to contribute to the potential of these pathogens to drive local and global
extinctions of some species [@Fisher2012; @Eskew2013]. Oomycetes, a taxonomically distinct 
group from fungi that possess many of these same traits, pose similar threats.

An important component of fungal host-pathogen dynamics is the role of spore load
in driving host infection and mortality. @Briggs2010 showed that
population-level persistence or extinction of mountain yellow-legged frogs (*Rana
sierrae*) infected by chytrid fungus (*Batrachochytrium dendrobatidis*), could
be explained by the dynamics of spore load build-up in these populations.
Similarly, mortality of bat populations
with White Nose Syndrome (*Geomyces destructans*) is closely related to spore
load on bat skin, which builds up through bat-to-bat contact over the course of
the hibernation period [@Langwig2015]. The fungal parasite *Metschnikowia
bicuspidata* kills its *Daphnia* host when the parasite loads are high enough to
interfere with cell metabolism [@Hall2009b]. Many mammals, including humans,
which are generally resistant to fungal diseases or asymptotic under normal
conditions may nonetheless become infected and exhibit symptoms or mortality
under persistent exposure to large fungal spore loads [@Casadevall2005].
Sudden Oak Death, caused by the oomycete *Phyophthora ramorum*,
kills tanoak (*Notholithocarpus densiflorus*) faster in the presence of large
numbers of other infected hosts, indicating that continued accumulation of
infections, rather than just disease progress, drives mortality [@Cobb2012].

Many attempts to model fungal disease dynamics have used traditional susceptible-infected (*SI*) disease frameworks, which represent
disease as a binary state of the host [@Kermack1927; @Briggs2005; @Mitchell2008; @Meentemeyer2011; @Foley2011a; @Maher2012;  @Cobb2012; @ORegan2014]
Extensions such as
susceptible-exposed-infected-removed (*SEIR*) models represent disease
progression within hosts, but do not capture the accumulation of new 
infections in hosts that may drive disease impacts. This dynamic may be better captured
using the framework created by @Anderson1978 to model macroparasite infections,
where disease is represented by the number of discrete infections (or
parasites), within each host. Here I use the term multi-infection to describe
these models.

Response to disease outbreak often requires prediction of medium- and long-term
behavior from early-phase dynamics of disease, which in turn requires
mechanistic disease models. However, in early stages of epidemics of emerging
diseases, the importance of disease load may not be known, nor the
appropriateness of multi-infection rather than *SI* models. Disease load is
often considerably more difficult to measure than prevalence, allowing only
observations of susceptible vs. infected states. However, model choice may have
considerable influence on predictions of disease dynamics and management
response.

Age and stage structure is another potentially important factor driving fungal
disease dynamics. The effect of Chytrid fungus on frogs varies across
life stages [@Rachowicz2004; @Garner2009]. Mortality rates in tanoak are much
greater in large trees than small ones [@Cobb2012]. However as organisms can
grow over similar time scales as the progress of a fungal epidemic, observed
patterns in age-disease relationships can be difficult to disentangle from
host-pathogen dynamics.

There is a considerable literature on host-pathogen dynamics of in age- and
stage-structured populations using *SI* models [@CastilloChavez1989; @Busenberg1990; @Diekmann1990; @Hethcote2000;
@Dietz2002; @Klepac2010]. There is a somewhat
smaller literature on age structure in multi-infection models. In
multi-infection models, @Krasnov2006 showed that parasite counts increase with
age in rodents. @Pacala1988 created a method to detect the mortality effect of
macroparasites based of the distribution of parasites among different age
groups. @Duerr2003 showed how age-infection relationships could be modified by
a variety of age, time, and density-dependent processes, but also showed that
interpretation of such age-infection patterns was ambiguous if more than one
such process was operating.

Here I explore how the transient host-pathogen dynamics differ between *SI* and
multi-infection systems. To compare models that with different structures and in
which parameters have different interpretations, I fit the models to each other
so that they have identical *dynamic behaviors* at different stages of an epidemic.
Using this approach, I ask, how do *SI* and multi-infection models that behave
similarly at early stages of disease differ in long-term dynamics, and vice versa? 
I also ask, how does model structure affect predictions of stage-specific behavior?
I find that mortality rates increase over time multi-infection models, diverging
from $SI$ models, and that these mortality rates diverge between host stages
due to accumulation of infections, even when stage-specific parameters are identical.
This suggests that the choice $SI$ or multi-infection as an appropriate framework to
represent fungal disease long-term disease predictions and interpretation of
stage-specific field data.


Methods
-------

### Model Structure

I compared dynamics in 3 ODE disease models: A simple $SI$ model, a
multi-infection model based on @Anderson1978, and an intermediate $SIV$
(susceptible-infected-*very* infected) model.

Each model has a two-stage population structure (population $N$ = juveniles
$J$ + adults $A$). New individuals enter the uninfected, juvenile stage via
density-dependent recruitment ($fN(1-N/K)$, where $f$ is fecundity and $K$
carrying capacity). Individuals move from juvenile to adult classes at the
transition rate $g$.

Disease transmission is density-dependent; In the *SI* model, susceptible individuals $(J_S, A_S)$
become infected $(J_I, A_I)$ at a rate equal to the density of infected
individuals times the transmissivity of the disease $(\lambda)$. All individuals
die at the a base rate $(d)$, and diseased individuals have additional mortality
$(\alpha)$.

The complete $SI$ model is

$$\begin{aligned}
  \frac{dJ_S}{dt} &= fN(1 - N/K) -
                     J_S(d + g + \lambda J_I + \lambda A_I)
  &\frac{dA_S}{dt} &= g J_S - A_S(d + \lambda N) \\
  \frac{dJ_I}{dt} &= \lambda J_S (J_I + A_I) - J_I(d + g + \alpha)
  &\frac{dA_I}{dt} &= g J_I + \lambda A_S (J_I + A_I) -
                      A_I(d + \alpha) \\
  N &= J_S + A_S + J_I + A_I
\end{aligned}$$

Note that in this model the effect of age structure is trivial: neither demographic nor
epidemiological parameters vary with age. When juvenile and adult classes are
summed, the growth term $g$ drops out, and $dN/dt$ is independent of $g$. Thus,
it is a "null model" for age effects.

The other two models are extensions of the $SI$ model with additional disease
classes representing degrees of infection. In the multi-infection model, there
are an infinite number of disease classes designated
$i = 0, 1, 2, \dots, \infty$. For purposes of simulation, the number of classes
is truncated, with a maximum value of $k$. Transmissivity $(\lambda)$ and
mortality $(\alpha)$ are additive in these models, increasing linearly with
$i$. Invividuals advance to the next disease class at rate $\Lambda$, the overall
force of infection, which is the sum of each tree's contribution, $i\lambda$.
Individuals in each stage die at rate $d + i\alpha$. Here is the complete
multi-infection model:

$$\begin{aligned}
  \frac{dJ_0}{dt} &= fN(1 - N/K) - J_0(d + g + \Lambda)
  &\frac{dA_0}{dt} &= g J_0 - A_0(d + \Lambda) \\
  \frac{dJ_i}{dt} &= \Lambda dJ_{i-1} -
                     J_i(d + g + i\alpha + \Lambda)
  &\frac{dA_i}{dt} &= g J_i + \Lambda A_{i-1} -
                     A_i(d + i\alpha + \Lambda) \\
  \frac{dJ_k}{dt} &= \Lambda dJ_{k-1} - J_k(d + g + k\alpha)
  &\frac{dA_k}{dt} &= g J_k + \Lambda A_{k-1} - A_k(d + k\alpha) \\
  N &= \sum_{i=0}^k J_0 + A_0
  &\Lambda &= \lambda \sum_{i=1}^k i(J_i + A_i)
\end{aligned}$$

The $SIV$ model is merely a truncated version of the multi-infection model, with
$k = 2$. For this model I refer $N_0$ as $S$, $N_1$ as $I$ and $N_2$ as $V$, and
use $S$, $I$, and $V$, as subscripts for $J$, and $A$ as well.

In this paper, parameters (e.g., $\lambda$ and $\alpha$) are subscripted with
$\text{param}_{SI}$, $\text{param}_{SIV}$, or $\text{param}_\text{multi}$ when
referring to their values in each of the three models. I also use the term
"infected" to refer to individuals of either the $I$ class in the $SI$ model, or
having at least one infection in the $SIV$ or multi-infection models.

Multi-infection models typically assume a distribution of infections in order to
reduce the infinite system of equations [@Anderson1978]. Negative-binomial distributions
of infections allow tractable analysis of such models and match empirical
studies of infection distribution in the wild [@Wilson2002]. However,  reduced models
only approximate the full model asymptotically and do not capture transient dynamics [@Adler1992], and the assumption of a negative-binomial distribution does not hold in the presence of age structure. (See Appendix A).
Instead, I avoided making such assumptions by simulating the the
infinite system of equations truncating at $k$. **I'd like more detail here -- you simulate $2k$ equations but encode a rule to increase $k$ can increase?**

### Comparative parameterization

I compared the models' behaviors under parameterizations determined "equivalent" by one of three of heuristic criteria described below. As the
models have different structures, parameters in the models have different
interpretations. Specifically, $\lambda$ and $\alpha$ operate on a
per-individual basis in the $SI$ model, while they operate on a per-infection
basis on the $SIV$ and multi-infection models. 

In order to determine equivalent parameterizations between models, I set
parameters for the $SI$ model to those in Table 1. I then fit the $SIV$ and
multi-infection models so that they would exhibit identical *behavior* to the
$SI$ model under different criteria. The behavior of SIV and multi-infection
models were adjusted by multiplying each of infectivity and disease-induced mortality
$(\lambda, \alpha)$ parameters by a constant, $c$.

Initial conditions in simulations were set at the disease-free equilibrium of
the system, modified with 1% of both juveniles and adults having a single
infection.

  Parameter                                          Symbol      Base Case Value
  ------------------------------------------------ ----------- -----------------
  fecundity                                            $f$                     1
  carrying capacity                                    $K$                     1
  transition rate                                      $g$                   0.1
  mortality                                            $d$                  0.01
  disease-induced mortality                         $\alpha$                 0.2
  transmissivity                                    $\lambda$                  3
  max number of infections (SIV/multi-infection)       $k$               3 / 150

Table 1: Base parameters for disease models

I examined model behavior in three cases. In each case using a different criterion to define equivalent parameterization:

1.  **Equilibrium mortality rate**. The first behavioral criterion was identical
    equilibrium mortality rate across models. $c$ was varied to match the
    overall disease-induced mortality rate (and thus the total mortality rate)
    between models. That is, at steady state,

    $$\alpha_{SI} = \alpha_{SIV} \frac{I + 2V}{I+V} 
              = \alpha_\text{multi} \frac{1}{N} \sum_i i N_i$$

2.  **Initial growth and acceleration rates of infected individuals.** Next, $c$
    was adjusted such that the second derivatives of growth of total
    infected individuals under initial conditions. (The first derivative is
    identical in all cases.) That is,

    $$\frac{d^2 I}{dt^2} = \frac{d^2(I+V)}{dt^2} = \frac{d^2 N_{i > 0}}{dt^2}$$
    at initial conditions of $S \approx N$,
    $I_{SI} = I_{SIV} = N_{1\, multi} \approx 0$ and
    $I_{SIV} = N_{i \geq 2\, multi} = 0$.

3.  **Time to 10% infection**. This criterion was selected to match behavior
    among models during the early transient period of disease. $c$ was adjusted
    so that the $SIV$ and multi-infection models would reach 10% infection in
    the same time period as the $SI$ model. That is,

    $$t\big|_{\frac{I}{S+I} = 0.1} = t\big|_{\frac{I+V}{S+I+V} = 0.1}
                            = t\big|_{\frac{N_{i \geq 1}}{N} = 0.1}$$

All simulations were performed in R [@RCORE], using the deSolve
package [@deSolve] for simulation, the numDeriv package to determine derivatives
[@numDeriv], and the ggplot2 [@ggplot2] package for plotting. Code to reproduce these
results is archived online [@Ross2014].

Results
-------

```{r base_models}
```

```{r modparms}
```

```{r modparms2}
```

```{r modparms3}
```

### Parameterizations

For the equilibrium parameterization, the constant $c$, or the ratio of $\alpha$ and $\lambda$ values between
the models, is the inverse of the mean number of infections at equilibrium
in the $SIV$ and multi-infection models. This is `r round(mod_SIV, 2)` for
the SIV model and `r round(mod_multi, 2)` for the multi-infection model. For the
acceleration-rate parameterization, $c$ is `r round(mod_SIV2, 2)` for both $SIV$ and multi-infection
models.  For the time-to-10% infection parameterization, $c$ was is `r round(mod_SIV3, 2)` for both models.
Results for both accelleration and time-to-10% infection parameterizations were
nearly identical (See Appendix), so I report only the equilibrium and time-to-10% results in
the sections below.

### Dynamics of mortality rates

**Models with similar mortality rates at invasion or equilibrium diverage at 
other times.**  Mortality rates of infected individuals in both the $SIV$ and 
multi-infection models evolve over time while they are constant the $SI$ model.
(Figure 1). In all cases, $SIV$ and multi-infection mortality rates increase over
time until equilibrium.  The increase in mortality is greater in the multi-infection
model than the $SIV$ model, though the pattern is qualitatively similar. 

```{r Fig2, fig.cap="Changes in mortality rates"}
<<mortrateplot>>
mort_rate_comp_plot
```


### Population effects

In the $SIV$ and multi-infection models, the apparent mortality rate of infected
individuals increases over time. Early in the epidemic, individuals have small
numbers of infections, thus the mortality rate across individuals with any level
of infection is low. As the epidemic progresses, the mean number of infections
per infected individual increases, raising the mortality rate of the infected
class until equilibrium is reached.

This change in mortality rates is driven by changes in the distribution of
infections over time (Figure 2). As the disease progresses
through the population in the $SIV$ model, the proportion of individuals in the
$I$ and $V$ classes increases for both juveniles and adults. Similarly, in the
multi-infection model, the mean number of infections in each individual
increases over time, increasing the mortality rate.

All three models, under all three parameterizations, exhibit some common
patterns in the dynamics of population stages. From the disease-free equilibrium
dominated by adults, disease outbreak decreases the population of adult stages
and and increases both the relative and absolute population of the juvenile
stages. The infected population of both stages increases, with the adult
infected stage reaching a peak before equilibrium and the juvenile infected
stage reaching a smaller equilibrium with no peak.


```{r Fig3, fig.cap="Dynamics of infection classes in the multi-infection model"}
progplot_multi
```

While equilibrium behaviors are identical and models start at the same initial
conditions, transient behavior differs. The time to equilibrium is greater in
the multi-infection model than the $SIV$ model, and greater in both than the
$SI$ model. It takes longer in the $SIV$ model, and longest in the
multi-infection model, for the disease to emerge.

```{r Fig1, fig.cap="Host population dynamics"}

<<popplot>>
population_comp_plot
```


**Models with similar initial mortality rate reach different equilibrium mortality rates.**
Figure 4 shows the dynamics of the three models in the case where the initial
first and second derivatives are equivalent. As in the matching-equilibrium
parameterizations, mortality for infected individuals increases over time until
equilibrium is reached. Unlike that parameterization, per-infection parameters
in the $SIV$/multi-infection models (mortality and infectivity), are similar to
per-individual parameters in the $SI$ model. $c$ is `r round(mod_SIV, 2)` for
the SIV model and `r round(mod_multi, 2)` for the multi-infection model. Total
mortality rates for the $SIV$ and multi-infection models start at the same
levels as the $SI$ model and diverge over time.


The $SIV$ and multi-infection models have nearly identical behavior. At
equilibrium, their populations are suppressed to lower levels than in the $SI$
model, and a smaller number of the individuals are infected. This is because the
difference in mortality rates of infected individuals between the $SI$ and the
other models is greater, increasing turnover of infected individuals. In this
case, the $SIV$ and multi-infection models reach equilibrium *before* the $SI$
model.



```{r Fig4, fig.cap="Dynamics of models parameterized to equivalent initial rates of growth of infectious individuals"}
<<init_derivs>>
init_derivs_comp_plot
```


```{r Fig5, fig.cap="Dynamics of models parameterized to equivalent time to 10% infected individuals"}
<<time_to_I10>>
time_to_I10_comp_plot
```

Patterns in the time-to-10%-infection parameterizations are
similar to the matched-second-derivative parameterization. Mortality rates for
individuals in $SIV$ and multi-infection models increase before reaching
equilibrium, resulting in lower population sizes and lower populations of
infected individuals at equilibrium. Dynamics for the $SIV$ and multi-infection
models are again very similar, though not as similar as in the
matched-derivative case. Also, in this case, the number of infected individuals
reaches a peak before going down to reach equilibrium levels.

### Age effects

**Multi-infection models generate age-dependent effects not found in SI
models.**



In the equivalent equilibrium mortality parameterization, The $SIV$ and
multi-infection stage structures are slower to reach an equilibrium than the
$SI$ stage structure, with the multi-infection case being slowest. This is
similar to the aggregate dynamics for this parameterization. In the case of
equivalent initial derivatives, as will as the case of equivalent time to 10%
infection, the change in age structure from the disease-free equilibrium is
greater in the $SIV$ and multi-infection models than the $SI$ model. at
equilibrium, there are more juveniles and fewer adults in the
$SIV$/multi-infection cases.

In the $SIV$ and multi-infection models, the mortality rate of infected
juveniles and adults increases as the disease progresses, and their mortality
rate diverges, with adults having greater mortality rates than juveniles at
equilibrium. This occurs in all parameterizations. The reason for this can be
found in figures 2 and 3, which show the distribution of infections for both
adult and juvenile populations over the course of the epidemic in $SIV$ and
multi-infection models. Adults and juveniles begin with equal mean numbers of
infections, but as the epidemic continues, adult trees accumulate more
infections than juveniles by both new infections on adult trees and
already-infected juveniles recruiting into the adult population.

In a multi-infection model with age structure, individuals accumulate infections
over time, resulting in more infections, and thus greater mortality and
infectivity, among older individuals than younger individuals. Even in the
absence of age-driven variation in how individuals respond to disease (that is,
in a "null model"), different behavior is observed between age groups. In an
$SI$ model, these differences do not arise.

<!-- ADD figs for:
1.  Models with equivalent equilibrium behavior will have very different
    invasion criteria. Multi-infection will be *less* likely to become
    epidemic from a single infection

2.  These effects increase as the transition rate from juvenile to adult age
    classes increase. this changes the relative amount of time individuals in
    each class have had to acquire infections.  -->

Discussion
----------

$SI$ and multi-infection models represent disease differently: as binary
states of an individual, or as accumulations of multiple infections across
individuals. As a result, they may produce different host-pathogen dynamics. The
choice of model structure has important consequences for the prediction of
host-pathogen disease dynamics.

Epidemics that appear to be well represented by $SI$ models during their
outbreak phase may no longer be well represented in later stages if they have
dynamics driven by multiple infections. A multi-infection model that behaves
like an $SI$ model in early stages will diverge from $SI$ behavior as increasing
infection loads result in greater per-individual mortality rates. I found
similar behavior regardless of the criterion used to determine early-phase
dynamics (derivatives or time-to-10%-infection). If the multi-infection system
system has an equilibrium (other than disease-free or extinction), the host
population will be lower at this equilibrium than in an $SI$ system due to
greater suppression from to disease, while the diseased population itself will
be lower as highly infected individuals suppressing population more. A *smaller*
fraction of the population will be infected at equilibrium, because of the short
survival of highly infected individuals.

** Clarify this sentence: state concretely what the similar patterns are**
**I found similar patterns in the reverse case, where early disease behavior was
predicted from equilibrium behavior using both $SI$ and multiple-infection
models.** Multiple-infection models with the same equilibrium behavior as $SI$
models are slower to emerge, as the small numbers of infections on initially
infected individuals transmit less disease and kill at slower rates than "fully"
infected individuals in the $SI$ model, which have higher per-individual
transmission mortality. The models converge when mean infection numbers in
individuals in the multi-infection model rise such that their virulence matches
individuals in the $SI$ model.

**I see the  phrase I've bolded below as encapsulating your most potentially interesting ecological finding. Is it novel?**

**Multiple-infection models indicate that age- or stage-related patterns in
disease mortality can arise from the accumulation of infections over time, even
in the absence of biological differences among age class in susceptibility to
disease.** Here I found that, in all parameterizations of multi-infection models,
adult mortality rates increased faster than juvenile mortality rates as
epidemics progressed, even though per-infection mortality rates were identical
between life stages. While in some fungal diseases, host-pathogen interactions
drive differences in virulence between life stages (e.g., chytrid fungus, see
@Rachowicz2004), this difference could explain part or all of stage-related
differences in mortality in other fungal diseases, such as in Sudden Oak Death
[@Cobb2012].

Multiple infection-models also showed increase population-level age effects of
disease. In simulations where both $SI$ and multi-infection models had similar
initial behavior, disease resulted in a shift from adult- to juvenile dominance
over the course of the epidemic, but in the multiple-infection model this shift
was greater, as adult mortality was greater. Also, adult disease prevalence was
lower in the late stages of multiple infection models, because with higher adult
mortality diseased adults have short lifespans.

Simplified models of multiple infection, such as the $SIV$ model presented here,
can capture some of the components of load-driven disease dynamics. Here, the
$SIV$ model behaved similarly to the multi-infection model, including similar
differences in time-to-equilibrium from the $SI$ model, similar suppression of
the final population, similar total infected host number, and age-mortality
patterns. However, there were differences between the $SIV$ and multi-infection
model in the apparent mortality rate of infected hosts, especially in the
time-to-10%-infection parameterization, indicating a role of the long tail of
hosts with high infection number in driving this pattern.

These results indicate that identifying multi-infection driven diseases early in
their emergence will **significantly** (Biologically significant?) alter predictions of disease dynamics. Can
the dynamics of these disease be distinguished from those of $SI$-like processes
in the data from early-stage emerging epidemics, especially when data are of
disease prevalence rather than load? One way to distinguish these mechanisms is
to look for changes in mortality rate as disease progresses or between early- or
late-epidemic populations, or to look for differences in mortality rate among age
classes. Both these patterns can indicate multi-infection-driven processes, though such patterns are not
sufficient to disentangle the multiple processes that may drive mortality
patterns. Instead, these patterns can indicate the potential role of these mechanisms,
and the need to investigate the relationships between infection load and host
effects.

<!-- 
Appendix [TODO]:
---------

1.  Show Klieber-Ludwig divergence of multi-infection model from negative
    binomial distribution during transient phase

2.  Show analysis of divergence of simplified version of multi-infection model
    [@Anderson1978] from full model when age structure.
-->

## References
