% Comparative Dynamics
% Noam Ross
% 14-08-25 10:10:34 

`r library(knitr); opts_chunk$set(echo=FALSE, cache=TRUE, message=FALSE, eval=FALSE)`

Here I examine the comparative dynamics of age classes in 2-age-class models
of both $SI$, $SIV$, and macroparasite-style dynamic models.

The following parameters are identical between models:

 - fecundity $(f)$
 - transition rate $(g)$
 - carrying capacity $(K)$
 
The models are also parameterized to have identical *apparent mortality* rates
at equilibrium.  That is the aggregate mortality rates of all infection classes
are the same. 

```{r setparms}
parms_SI = list( 
  n_stages = 2,
  n_parasites = 1,
  f = c(1, 1),
  g = c(0.1, 0),
  d = c(0.01, 0.01),
  alpha = c(0.2, 0.2),
  lamda = c(0.5, 0.5),
  Beta = c(1, 1),
  mu = c(0, 0),
  xi = c(1, 1),
  omega = c(1,1),
  K = 1,
  progress = 0,
  times.min = 0,
  times.max = 50,
  times.by = 0.1,
  inits = 0,
  infect_vector = c(0.01, 0.01)
  )

parms_SIV =  parms_SI
parms_SIV$n_parasites = 2

parms_macro = parms_SI
parms_macro$n_parasites = 150
parms_SI$times.by = 0.01
```

```{r setup}
library(deSolve)
library(plyr)
library(rootSolve)
library(reshape2)
library(ggplot2)
library(data.table)
library(noamtools)
library(numDeriv)
source('R/process_sodp.R')
source('R/run_sodp.R')
```

```{r runmodels}
out_SI = run_sodp(parms_SI)
SIsf = process_sodp(out_SI)
SIsf = sodp_totals(SIsf)
stats_SI = sodp_stats(SIsf)
stats_SI.m = melt(stats_SI, id.vars=c("time", "Species", "SizeClass"))

out_SIV = run_sodp(parms_SIV)
SIVsf = process_sodp(out_SIV)
SIVsf = sodp_totals(SIVsf)
stats_SIV = sodp_stats(SIVsf)
stats_SIV.m = melt(stats_SIV, id.vars=c("time", "Species", "SizeClass"))

out_macro = run_sodp(parms_macro)
macrosf = process_sodp(out_macro)
macrosf = sodp_totals(macrosf)
stats_macro = sodp_stats(macrosf)
stats_macro.m = melt(stats_macro, id.vars=c("time", "Species", "SizeClass"))

```


```{r initplots}
ggplot(stats_SI.m, aes(x=time, y=value, col=SizeClass)) + 
 facet_wrap(~variable, scales="free_y") +
 geom_line() 

ggplot(stats_SIV.m, aes(x=time, y=value, col=SizeClass)) + 
 facet_wrap(~variable, scales="free_y") +
 geom_line() 

ggplot(stats_macro.m, aes(x=time, y=value, col=SizeClass)) + 
 facet_wrap(~variable, scales="free_y") +
 geom_line() 
```

```{r complots}
stats = list(SI=stats_SI.m, SIV=stats_SIV.m, macro=stats_macro.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="MortInfRate"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass, size=SizeClass)) +
  geom_line() +
  ylim(0, 0.4) +
  ylab("Mortality Rate of Infected Trees") +
  xlab("Time") +
  scale_linetype_manual(labels=c("Juv", "Adult", "Total"), values=c(3,2,1)) +
<<<<<<< HEAD
  theme_nr +
  theme(text=element_text(family="Lato")) +
  guides(linetype = guide_legend(reverse=TRUE))

=======
  scale_size_manual(values=c(0.5,0.5, 1.5), guide='none') + 
  theme_nr +
  theme(text=element_text(family="Lato")) +
  guides(linetype = guide_legend(reverse=TRUE))
>>>>>>> 30628c54b6b22941fc2ff6fa7dd30076b06b3ea5
```

```{r modparms}

parms_SIVmod = parms_SIV



aa = optim(par = 1, f = function(mod_SIV) {
  parms_SIVmod$alpha = parms_SIV$alpha*mod_SIV
  parms_SIVmod$lamda = parms_SIV$lamda*mod_SIV
  out_SIVmod = run_sodp(parms_SIVmod)
  SIVmodsf = process_sodp(out_SIVmod)
  SIVmodsf = sodp_totals(SIVmodsf)
  stats_SIVmod = sodp_stats(SIVmodsf)
  val = subset(stats_SIVmod, time==parms_SIVmod$times.max & SizeClass=="Total")$MortInfRate
  if(interactive()) print(val)
  return((val - parms_SI$alpha[1])^2)
  }, method="Brent", lower=0, upper=2)


mod_SIV = aa$par
parms_SIVmod$alpha = parms_SIV$alpha*mod_SIV
parms_SIVmod$lamda = parms_SIV$lamda*mod_SIV
out_SIVmod = run_sodp(parms_SIVmod)
SIVmodsf = process_sodp(out_SIVmod)
SIVmodsf = sodp_totals(SIVmodsf)
stats_SIVmod = sodp_stats(SIVmodsf)
stats_SIVmod.m = melt(stats_SIVmod, id.vars=c("time", "Species", "SizeClass"))


parms_macromod = parms_macro


bb = optim(par = 0.6118, f = function(mod_macro) {
  parms_macromod$alpha = parms_macro$alpha*mod_macro
  parms_macromod$lamda = parms_macro$lamda*mod_macro
  out_macromod = run_sodp(parms_macromod)
  macromodsf = process_sodp(out_macromod)
  macromodsf = sodp_totals(macromodsf)
  stats_macromod = sodp_stats(macromodsf)
  val = subset(stats_macromod, time==parms_macromod$times.max & SizeClass=="Total")$MortInfRate
  if(interactive()) print(val)
  return((val - parms_SI$alpha[1])^2)
  }, method="Brent", lower=0, upper=2)

mod_macro = bb$par

parms_macromod$alpha = parms_macro$alpha*mod_macro
parms_macromod$lamda = parms_macro$lamda*mod_macro

out_macromod = run_sodp(parms_macromod)
macromodsf = process_sodp(out_macromod)
macromodsf = sodp_totals(macromodsf)
stats_macromod = sodp_stats(macromodsf)
stats_macromod.m = melt(stats_macromod, id.vars=c("time", "Species", "SizeClass"))

# ggplot(stats_SI.m, aes(x=time, y=value, col=SizeClass)) + 
#  facet_wrap(~variable, scales="free_y") +
#  geom_line() 
# 
# 
# ggplot(stats_SIVmod.m, aes(x=time, y=value, col=SizeClass)) + 
#  facet_wrap(~variable, scales="free_y") +
#  geom_line() 
# 
# ggplot(stats_macromod.m, aes(x=time, y=value, col=SizeClass)) + 
#  facet_wrap(~variable, scales="free_y") +
#  geom_line() 
```

```{r plotcomp}
stats = list(SI=stats_SI.m, SIV=stats_SIVmod.m, macro=stats_macromod.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="MortInfRate"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
  ylim(0, 0.4) +
  ylab("Mortality Rate of Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

stats = list(SI=stats_SI.m, SIV=stats_SIVmod.m, macro=stats_macromod.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="N"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Total Population") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

stats = list(SI=stats_SI.m, SIV=stats_SIVmod.m, macro=stats_macromod.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="I"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

# ggplot(subset(statsall, variable %in% c("N", "P", "S", "I", "FracInf", "Mean", "MortInfRate") & time <= 100),
#        aes(x=time, y=value, col=SizeClass)) + 
#  facet_grid(.id~variable) +
#  geom_line() 
```

```{r modparms2}

SI_Final = init_derivs2(parms_SI, SI=TRUE)[,2]

parms_SIVmod2 = parms_SIV



aa = optim(par = c(1,1), fn = function(mod_SIV) {
  parms_SIVmod$alpha = parms_SIV$alpha*mod_SIV[1]
  parms_SIVmod$lamda = parms_SIV$lamda*mod_SIV[2]
  val = init_derivs2(parms_SIVmod, SI=TRUE)[,2]
  if(interactive()) print(val)
  return(sum(((val - SI_Final)/SI_Final)^2))
  })


mod_SIV = aa$par
parms_SIVmod2$alpha = parms_SIV$alpha*mod_SIV[1]
parms_SIVmod2$lamda = parms_SIV$lamda*mod_SIV[2]
out_SIVmod2 = run_sodp(parms_SIVmod2)
SIVmod2sf = process_sodp(out_SIVmod2)
SIVmod2sf = sodp_totals(SIVmod2sf)
stats_SIVmod2 = sodp_stats(SIVmod2sf)
stats_SIVmod2.m = melt(stats_SIVmod2, id.vars=c("time", "Species", "SizeClass"))


parms_macromod2 = parms_macro


bb = optim(par = aa$par, f = function(mod_macro) {
  parms_macromod$alpha = parms_macro$alpha*mod_macro[1]
  parms_macromod$lamda = parms_macro$lamda*mod_macro[2]
  val = init_derivs2(parms_macromod, SI=TRUE)[,2]
  if(interactive()) print(val)
  return(sum(((val - SI_Final)/SI_Final)^2))
  })

mod_macro = bb$par

parms_macromod2$alpha = parms_macro$alpha*mod_macro[1]
parms_macromod2$lamda = parms_macro$lamda*mod_macro[2]

out_macromod2 = run_sodp(parms_macromod2)
macromod2sf = process_sodp(out_macromod2)
macromod2sf = sodp_totals(macromod2sf)
stats_macromod2 = sodp_stats(macromod2sf)
stats_macromod2.m = melt(stats_macromod2, id.vars=c("time", "Species", "SizeClass"))

```

```{r plotcomp2}
stats = list(SI=stats_SI.m, SIV=stats_SIVmod2.m, macro=stats_macromod2.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="MortInfRate"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
 # ylim(0, 0.4) +
  ylab("Mortality Rate of Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

stats = list(SI=stats_SI.m, SIV=stats_SIVmod2.m, macro=stats_macromod2.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="N"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Total Population") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

stats = list(SI=stats_SI.m, SIV=stats_SIVmod2.m, macro=stats_macromod2.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="FracInf"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))
```


```{r modparms3}


frac_I = 0.1
sit = subset(stats_SI, SizeClass=="Total", select=c("time", "FracInf"))
sit = sit[order(sit$time),]
timeToFracI_fn = function(sit, frac_I) {
  afti = min(which(sit$FracInf >= frac_I), na.rm=TRUE)
  if(is.na(afti) || afti==Inf) return(max(sit$time))
  befi = min(which(sit$FracInf >= frac_I)) - 1
  befx = sit$time[befi]
  befy = sit$FracInf[befi]
  aftx = sit$time[afti]
  afty = sit$FracInf[afti]
  slope = (afty-befy)/(aftx-befx)
  return(befx + (frac_I - befy)/slope)
  }
timeToFracI = timeToFracI_fn(sit, frac_I)
parms_SIVmod3 = parms_SIV
parms_SIVmod3$times.max = 18
parms_SIVmod3$times.by = 0.01


cc = optim(par = 1, fn = function(mod_SIV) {
  parms_SIVmod3$alpha = parms_SIV$alpha*mod_SIV
  parms_SIVmod3$lamda = parms_SIV$lamda*mod_SIV
  out_SIVmod = run_sodp(parms_SIVmod3)
  SIVmodsf = process_sodp(out_SIVmod)
  SIVmodsf = sodp_totals(SIVmodsf)
  stats_SIVmod = sodp_stats(SIVmodsf)
  sitSIVmod = subset(stats_SIVmod, SizeClass=="Total", select=c("time", "FracInf"))
  sitSIVmod = sitSIVmod[order(sitSIVmod$time),]
  timeToFracISIVmod = timeToFracI_fn(sitSIVmod, frac_I)
  if(interactive()) print(timeToFracISIVmod)
  val = ((timeToFracISIVmod - timeToFracI)/timeToFracI)^2
  return(val)
  }, method="Brent", lower=0, upper=2)


mod_SIV = cc$par
parms_SIVmod3$alpha = parms_SIV$alpha*mod_SIV
parms_SIVmod3$lamda = parms_SIV$lamda*mod_SIV
parms_SIVmod3$times.max = 50
parms_SIVmod3$times.by = 0.1
out_SIVmod3 = run_sodp(parms_SIVmod3)
SIVmod3sf = process_sodp(out_SIVmod3)
SIVmod3sf = sodp_totals(SIVmod3sf)
stats_SIVmod3 = sodp_stats(SIVmod3sf, dofit=FALSE)
stats_SIVmod3.m = melt(stats_SIVmod3, id.vars=c("time", "Species", "SizeClass"))


parms_macromod3 = parms_macro
parms_macromod3$times.max = 18
parms_macromod3$times.by = 0.01

dd = optim(par = cc$par, f = function(mod_macro) {
  parms_macromod3$alpha = parms_macro$alpha*mod_macro
  parms_macromod3$lamda = parms_macro$lamda*mod_macro
  out_macromod = run_sodp(parms_macromod3)
  macromodsf = process_sodp(out_macromod)
  macromodsf = sodp_totals(macromodsf)
  stats_macromod = sodp_stats(macromodsf)
  sitmacromod = subset(stats_macromod, SizeClass=="Total", select=c("time", "FracInf"))
  sitmacromod = sitmacromod[order(sitmacromod$time),]
  timeToFracImacromod = timeToFracI_fn(sitmacromod, frac_I)
  if(interactive()) print(timeToFracImacromod)
  val = ((timeToFracImacromod - timeToFracI)/timeToFracI)^2
  return(val)
  }, method="Brent", lower=0, upper=2)

mod_macro = dd$par

parms_macromod3$alpha = parms_macro$alpha*mod_macro
parms_macromod3$lamda = parms_macro$lamda*mod_macro
parms_macromod3$times.max = 50
parms_macromod3$times.by = 0.1
out_macromod3 = run_sodp(parms_macromod3)
macromod3sf = process_sodp(out_macromod3)
macromod3sf = sodp_totals(macromod3sf)
stats_macromod3 = sodp_stats(macromod3sf, dofit=FALSE)
stats_macromod3.m = melt(stats_macromod3, id.vars=c("time", "Species", "SizeClass"))

```

```{r plotcomp3}
stats = list(SI=stats_SI.m, SIV=stats_SIVmod3.m, macro=stats_macromod3.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="MortInfRate"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
 # ylim(0, 0.4) +
  ylab("Mortality Rate of Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

stats = list(SI=stats_SI.m, SIV=stats_SIVmod3.m, macro=stats_macromod3.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="N"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Total Population") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

stats = list(SI=stats_SI.m, SIV=stats_SIVmod3.m, macro=stats_macromod3.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="FracInf"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Fraction Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))
```

```{r modparms4}


frac_I = 0.1
frac_I2 = 0.3
sit = subset(stats_SI, SizeClass=="Total", select=c("time", "FracInf"))
sit = sit[order(sit$time),]
timeToFracI = timeToFracI_fn(sit, frac_I)
timeToFracI[2] = timeToFracI_fn(sit, frac_I2)


parms_SIVmod4 = parms_SIV
parms_SIVmod4$progress = 0
parms_SIVmod4$times.max = 25
parms_SIVmod4$times.by = 0.01

ee = optim(par = c(1, 1), fn = function(mod_SIV) {
  parms_SIVmod4$alpha = parms_SIV$alpha*mod_SIV[1]
  parms_SIVmod4$lamda = parms_SIV$lamda*mod_SIV[2]
  out_SIVmod = run_sodp(parms_SIVmod4)
  SIVmodsf = process_sodp(out_SIVmod)
  SIVmodsf = sodp_totals(SIVmodsf)
  stats_SIVmod = sodp_stats(SIVmodsf)
  sitSIVmod = subset(stats_SIVmod, SizeClass=="Total", select=c("time", "FracInf"))
  sitSIVmod = sitSIVmod[order(sitSIVmod$time),]
  timeToFracISIVmod = timeToFracI_fn(sitSIVmod, frac_I)
  timeToFracISIVmod[2] = timeToFracI_fn(sitSIVmod, frac_I2)
  if(interactive()) print(timeToFracISIVmod)
  val = sum(((timeToFracISIVmod - timeToFracI)/timeToFracI)^2)
  return(val)
  })


mod_SIV = ee$par
parms_SIVmod4$alpha = parms_SIV$alpha*mod_SIV[1]
parms_SIVmod4$lamda = parms_SIV$lamda*mod_SIV[2]
parms_SIVmod4$times.max = 50
parms_SIVmod4$times.by = 0.1
out_SIVmod4 = run_sodp(parms_SIVmod4)
SIVmod4sf = process_sodp(out_SIVmod4)
SIVmod4sf = sodp_totals(SIVmod4sf)
stats_SIVmod4 = sodp_stats(SIVmod4sf, dofit=FALSE)
stats_SIVmod4.m = melt(stats_SIVmod4, id.vars=c("time", "Species", "SizeClass"))


parms_macromod4 = parms_macro
parms_macromod4$progress = 0
parms_macromod4$times.max = 25
parms_macromod4$times.by = 0.01


ff = optim(par = ee$par, f = function(mod_macro) {
  parms_macromod4$alpha = parms_macro$alpha*mod_macro[1]
  parms_macromod4$lamda = parms_macro$lamda*mod_macro[2]
  out_macromod = run_sodp(parms_macromod4)
  macromodsf = process_sodp(out_macromod)
  macromodsf = sodp_totals(macromodsf)
  stats_macromod = sodp_stats(macromodsf)
  sitmacromod = subset(stats_macromod, SizeClass=="Total", select=c("time", "FracInf"))
  sitmacromod = sitmacromod[order(sitmacromod$time),]
  timeToFracImacromod = timeToFracI_fn(sitmacromod, frac_I)
  timeToFracImacromod[2] = timeToFracI_fn(sitmacromod, frac_I2)
  if(interactive()) print(timeToFracImacromod)
  val = sum(((timeToFracImacromod - timeToFracI)/timeToFracI)^2)
  return(val)
  })

mod_macro = ff$par

parms_macromod4$alpha = parms_macro$alpha*mod_macro[1]
parms_macromod4$lamda = parms_macro$lamda*mod_macro[2]
parms_macromod4$times.max = 50
parms_macromod4$times.by = 0.1
out_macromod4 = run_sodp(parms_macromod4)
macromod4sf = process_sodp(out_macromod4)
macromod4sf = sodp_totals(macromod4sf)
stats_macromod4 = sodp_stats(macromod4sf, dofit=FALSE)
stats_macromod4.m = melt(stats_macromod4, id.vars=c("time", "Species", "SizeClass"))

```

```{r plotcomp4}
stats = list(SI=stats_SI.m, SIV=stats_SIVmod4.m, macro=stats_macromod4.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="MortInfRate"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
 # ylim(0, 0.4) +
  ylab("Mortality Rate of Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))


mortInf = ldply(stats, function(x) subset(x, variable=="N"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Total Population") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

mortInf = ldply(stats, function(x) subset(x, variable=="FracInf"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Fraction Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

mortInf = ldply(stats, function(x) subset(x, variable=="I"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

```


```{r modparms5}


frac_I = 0.1
sit = subset(stats_SI, SizeClass=="Total", select=c("time", "FracInf"))
sit = sit[order(sit$time),]
timeToFracI = timeToFracI_fn(sit, frac_I)
timeToFracI[2] = subset(stats_SI, time==parms_SI$times.max & SizeClass=="Total")$FracInf

parms_SIVmod5 = parms_SIV
parms_SIVmod5$progress = 0
parms_SIVmod5$times.by = 0.01

gg = optim(par = c(1, 1), fn = function(mod_SIV) {
  parms_SIVmod5$alpha = parms_SIV$alpha*mod_SIV[1]
  parms_SIVmod5$lamda = parms_SIV$lamda*mod_SIV[2]
  out_SIVmod = run_sodp(parms_SIVmod5)
  SIVmodsf = process_sodp(out_SIVmod)
  SIVmodsf = sodp_totals(SIVmodsf)
  stats_SIVmod = sodp_stats(SIVmodsf)
  sitSIVmod = subset(stats_SIVmod, SizeClass=="Total", select=c("time", "FracInf"))
  sitSIVmod = sitSIVmod[order(sitSIVmod$time),]
  timeToFracISIVmod = timeToFracI_fn(sitSIVmod, frac_I)
  timeToFracISIVmod[2]= subset(stats_SIVmod, time==parms_SIVmod5$times.max & SizeClass=="Total")$FracInf
  if(interactive()) print(timeToFracISIVmod)
  val = sum(((timeToFracISIVmod - timeToFracI)/timeToFracI)^2)
  return(val)
  })


mod_SIV = gg$par
parms_SIVmod5$alpha = parms_SIV$alpha*mod_SIV[1]
parms_SIVmod5$lamda = parms_SIV$lamda*mod_SIV[2]
parms_SIVmod5$times.by = 0.1
out_SIVmod5 = run_sodp(parms_SIVmod5)
SIVmod5sf = process_sodp(out_SIVmod5)
SIVmod5sf = sodp_totals(SIVmod5sf)
stats_SIVmod5 = sodp_stats(SIVmod5sf, dofit=FALSE)
stats_SIVmod5.m = melt(stats_SIVmod5, id.vars=c("time", "Species", "SizeClass"))


parms_macromod5 = parms_macro
parms_macromod5$progress = 0
parms_macromod5$times.by = 0.01


hh = optim(par = gg$par, f = function(mod_macro) {
  parms_macromod5$alpha = parms_macro$alpha*mod_macro[1]
  parms_macromod5$lamda = parms_macro$lamda*mod_macro[2]
  out_macromod = run_sodp(parms_macromod5)
  macromodsf = process_sodp(out_macromod)
  macromodsf = sodp_totals(macromodsf)
  stats_macromod = sodp_stats(macromodsf)
  sitmacromod = subset(stats_macromod, SizeClass=="Total", select=c("time", "FracInf"))
  sitmacromod = sitmacromod[order(sitmacromod$time),]
  timeToFracImacromod = timeToFracI_fn(sitmacromod, frac_I)
  timeToFracImacromod[2] = subset(stats_macromod, time==parms_macromod5$times.max & SizeClass=="Total")$FracInf
  if(interactive()) print(timeToFracImacromod)
  val = sum(((timeToFracImacromod - timeToFracI)/timeToFracI)^2)
  return(val)
  })

mod_macro = hh$par

parms_macromod5$alpha = parms_macro$alpha*mod_macro[1]
parms_macromod5$lamda = parms_macro$lamda*mod_macro[2]
parms_macromod5$times.max = 50
parms_macromod5$times.by = 0.1
out_macromod5 = run_sodp(parms_macromod5)
macromod5sf = process_sodp(out_macromod5)
macromod5sf = sodp_totals(macromod5sf)
stats_macromod5 = sodp_stats(macromod5sf, dofit=FALSE)
stats_macromod5.m = melt(stats_macromod5, id.vars=c("time", "Species", "SizeClass"))

```

```{r plotcomp4}
stats = list(SI=stats_SI.m, SIV=stats_SIVmod5.m, macro=stats_macromod5.m)
statsall = ldply(stats, function(x) x)
mortInf = ldply(stats, function(x) subset(x, variable=="MortInfRate"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
 # ylim(0, 0.4) +
  ylab("Mortality Rate of Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))


mortInf = ldply(stats, function(x) subset(x, variable=="N"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Total Population") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))


mortInf = ldply(stats, function(x) subset(x, variable=="I"))
ggplot(mortInf, aes(x = time, y=value, col=.id, linetype=SizeClass)) +
  geom_line() +
#  ylim(0, 0.4) +
  ylab("Infected Trees") +
  xlab("Time") +
  scale_linetype(labels=c("Juv", "Adult", "Total")) +
  theme_nr +
  theme(text=element_text(family="Lato"))

```

Currently matching model by equilibrium mortality.  Next, match the models by
initial growth behavior.  That is, growth rate at time zero, or R0.   How do
models behave comparatively if they have initial growth rate of infected
individuals that is the same?  What if we peg both initial rate and final
equilibrium mortality rate, but free both $\alpha$ and $\lambda$ to vary
independently?  How will models behave differently?

Need too calculate this rate for the given initial number of infected individuals
Model may need to be modified to only have initial infected individuals be of
lowest disease class, not all disease classes.

At initiation, when all individuals are either susceptible or only have one
infection, the growth in infected individuals is

$$\lambda_{I_1} - d - \alpha$$


Note that the 2nd derivative of $V$ is very small no matter when $I$ is small at $V=H=0$:

$$\frac{d^2V}{dt^2} = - I^{3} \lambda^{2} + I^{2} \left(2 S \lambda^{2} - 4 \alpha \lambda - 3 d \lambda\right)$$